\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-11-09 k-planar.tex
Touched up figure

\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}. The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.  To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}. Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ that contains a tree rooted at each vertex of $V_0$ with the property that for each $v\in V(G)$,
\[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]

@@

For readers already familiar with the proof in \cite{dujmovic.joret.ea:planar}, the main difference here is that a typical tripod in the proof of \cite[Lemma~17]{dujmovic.joret.ea:planar} consists of three upward paths in a BFS tree/forest whose lower endpoints belong to a single triangular face $\tau$.  In the current setting, a typical tripod consists of three upward paths whose lower endpoints belong to a common face $\tau$ of $G_0$ plus all of the (at most $d$) vertices of $\tau$ that are not already included in tripods.

>>>>

The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A $T$-decomposition of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags contain $v_1,\ldots,v_t$.  To show the planarity of $H$ in \cref{d_framed_product_stucture} we use the fact that a graph has a 3-simple tree-decomposition if and only if it has treewidth at most 3 and is planar~\citep{kratochvil.vaner:note}.

@@

A \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ that contains a tree rooted at each vertex of $V_0$ with the property that for each $v\in V(G)$,
\[  \min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}.  \]
\cref{d_framed_product_stucture} is a consequence of the following technical lemma, which is an extension of the analogous result for plane graphs~\cite[Lemma~17]{DJMMUW20}.  For readers already familiar with the proof in \cite{dujmovic.joret.ea:planar}, the main difference here is that a typical tripod in the proof of \cite[Lemma~17]{dujmovic.joret.ea:planar} consists of three upward paths in a BFS tree/forest whose lower endpoints belong to a single triangular face $\tau$.  In the current setting, a typical tripod consists of three upward paths whose lower endpoints belong to a common face $\tau$ of $G_0$ plus all of the (at most $d$) vertices of $\tau$ that are not already included in tripods.
